---
title: "Plot assignment"
author: "Template: Ian Hussey; content: David Schlegel"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r, include=FALSE}

knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)

```

# Dependencies

```{r}
library(tidyverse)
library(janitor) 
library(openxlsx)
library(ggplot2)
library(introdataviz)
library(viridis)
library(ggsignif)
library(ggimage)

```

# Instructions

The goal of this exercise is to produce both good and bad examples of plots *and to be explicit about what makes them good or bad*. 

The data can be real or fabricated. The plots can convey realistic or silly messages - but they must convey them clearly. You can get real data from many sources. A good one is the open science framekwork (osf.io). You can also simply make up data, either by hand or using simulation functions (e.g., `rnorm`) or the many simulation packages. If simulating data, ensure that you (a) use set.seed() to make it reproudbile, and (b) save a copy of the final data that you're using to disk and then reload it in the script so that the plot is reproducible.

The goal is neither to reproduce the types of plots that you most often see in journals (bar plots, etc.) nor to use a very fancy looking plot just for the sake of it - if it is very unlikely that you'd ever be able to use the plot to convey a research finding in an article, blog, poster, tweet, infographic or otherwise, it's probably not a good choice either.

The plots must be reproducible. Provide the data as well as the code to reporduce them. The plots should also be saved as individual images (pdf or png).

Under each plot, here in the RMarkdown, briefly explain what types of analysis the plot is suitable for, what elements of the results the plot attempts to clearly convey. For the bad plot, do the opposite: emphasise what features of the results the plot obscures or makes unclear. In doing so, communicate to me that you understand what makes a good plot to convey the results of quantiative analyses to viewers.

Consider making use of the many extensions to ggplot2, e.g.:

- https://r-graph-gallery.com
- https://krzjoa.github.io/awesome-r-dataviz/#/
- ggdist https://mjskay.github.io/ggdist/
- ggrepel
- ggpubr
- see
- ggsignif
- and others: https://exts.ggplot2.tidyverse.org/gallery/

- osf.io
- AIID dataset: https://osf.io/pcjwf/


# Daten einlesen
```{r}
data_raw <- read_csv("../data/processed/depression_data.csv") |> 
  janitor::clean_names()
```

# Good plot

```{r}
rain_height <- 0.15

p1 <- ggplot(data_raw, aes(x = gender, y = depression_level, fill = gender)) +
      geom_flat_violin(trim = FALSE, alpha = 0.4,
                       adjust = 0.8, 
                       position = position_nudge(x = 0.15), 
                       size = 0.3) +
      geom_point(aes(color = stress_level), size = 3, alpha = .7, 
                 position = position_jitter(width = 0.1, height = 0)) +
      scale_color_gradient(low = "blue", high = "red", 
                           breaks = c(1,2.5,5,7.5,10), 
                           labels = c("1 - no stress", "2.5", "5 - medium stress", 
                           "7.5", "10 - high stress")) +
      # scale_color_viridis_c(option = "inferno", begin = 0.3, end = 0.9, direction = -1,
      #                       breaks = c(1,2.5,5,7.5,10)) +
      geom_boxplot(width = 0.1, alpha = 0.4, outlier.shape = NA,
                   position = position_nudge(x = -0.2), size = 0.3) +
      geom_signif(annotation = "p < 0.001",
                  xmin = 0.8, 
                  xmax = 1.8,
                  y_position = 13.5,
                  tip_length = 0.03, 
                  size = 0.3, 
                  textsize = 4) + 
      coord_flip() +
      labs(title = "Gender Differences in Depression Severity",
           subtitle = "Impact of Stress Levels on Depression Severity Among Men and Women",
           x = "Gender",
           y = "Depression Severity",
           color = "Stress Level") + 
      theme_minimal() +
      scale_fill_viridis_d(option = "mako", begin = 0.4, end = 0.7, direction = -1) +
      guides(fill = FALSE, color = guide_legend(title.position = "top", title.hjust = 0.5)) +
      theme(legend.position = "right",
           legend.background = element_rect(fill = "white", colour = "black", size = 0.1),
           legend.title = element_text(size = 13),
           legend.text = element_text(size = 12),
           axis.text.y = element_text(size = 14, face = "bold"), # Male / Female
           axis.text.x = element_text(size = 13), # Scale Deprecivity
           axis.title.x = element_text(size = 15), # Depression Severity
           axis.title.y = element_text(size = 15), # Gender
           plot.title = element_text(size = 19,face = "bold"),
           plot.subtitle = element_text(size = 16)) +
      scale_x_discrete(labels = c("female" = "Female", "male" = "Male")) +
      scale_y_continuous(breaks = c(1, 5, 10))

```

This plot, known as a rainplot, combines elements of a violin plot, box plot, and scatter plot, offering a multi-layered approach to visualizing continuous variables.

The violin plot component allows for comparison of the depression severity score distributions across gender groups. It visually represents the data density and indicates where scores are most concentrated, providing a clear picture of the distribution shapes.

The box plot offers a summary of the central tendencies for each gender category, with the medians highlighted, signaling any differences between the groups.

The scatter plot layer displays individual depression severity scores, here color-coded by stress level, facilitating the identification of patterns or outliers. Specifically, it enables an investigation into the relationship between stress and depression severity.

An additional analytical layer indicates statistically significant differences between the groups, adding depth to the interpretation of the data.

The color scheme is intentionally selected to be easily distinguishable and to ensure that colors representing stress levels are distinct, avoiding any potential confusion. The stress level colors transition from 'blue' for lower stress to 'red' for higher stress, aiding in quick visual differentiation.

Labels and legends are designed for clarity and readability. The layout maximizes space utilization, and the legend is placed to ensure good readabilty.

In essence, this plot strives to effectively communicate the distribution and central tendencies of depression severity across gender while examining the interplay between depression severity and stress levels. 

# Bad plot

```{r}

p2 <- ggplot(data_raw, aes(x = age, y = depression_level, color = therapy_type)) +
      geom_point(aes(size = stress_level), alpha = 0.5) +
      geom_smooth(method = "loess", se = FALSE) +
      facet_grid(gender ~ education_level) +
      scale_color_manual(values = c("psychoanalysis" = "yellow", "behavioral therapy" = "yellow3", "medication" = "violet", "no therapy" = "pink")) +
      theme_minimal() +
      labs(title = "",
           x = "",
           y = "Depression",
           color = "",
           size = "Stress") +
      theme(legend.position = "top",
            legend.direction = "vertical",
            legend.title = element_text(size = 15),
            legend.text = element_text(size = 11),
            axis.text.y = element_text(size = 12), 
            axis.text.x = element_text(size = 12), 
            axis.title.x = element_text(size = 9), 
            axis.title.y = element_text(size = 8)) 


```
The plot lacks a clear title to immediately inform the viewer of the purpose or the main message of the visualization. Additionally, the labels are either missing or unclear. 

The legend titled "Stress" uses circles of varying sizes to represent data points, but without a clear scale or explanation, it's difficult for the viewer to interpret the significance of the size differences accurately. The legend is on the top and the space is not used efficiently. 

The colors used to represent different therapy conditions are too similar, especially the yellow hues, which can easily be confused with one another. This lack of contrast makes it challenging to distinguish between therapy types.

The significant overlap of data points, particularly noticeable in certain therapy categories, obscures individual data and can lead to misinterpretation of the data distribution or trends.

The plot includes multiple trend lines without a legend explaining what they represent. The excess of lines clutters the plot and complicates the interpretation.

The plot tries to convey too many relationships and data points simultaneously, leading to a crowded and confusing visual. Without a clear focus, it's challenging to understand what the key takeaways should be.

# Write to disk

Write to disk as a .pdf or .png.

Name them "plot_good_[yourname]" and "plot_bad_[yourname]".

```{r}

ggsave("../communication/plot_good_David_Schlegel.png", p1, bg = "white", 
       width = 10, height = 8, dpi = 300)

ggsave("../communication/plot_bad_David_Schlegel.png", p2, bg = "white")

```

# Session info

```{r}

sessionInfo()

```


