---
title: "Plot Assignment"
subtitle: "Data processing"
author: "David Schlegel"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r, include=FALSE}

knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)

```

# Dependencies

```{r}

library(tidyverse)
library(janitor) 
library(stringr)
library(openxlsx)

```

# Get data

```{r}

# demographics
data_raw <- read_csv2("../data/raw/data_raw.csv", ) |> 
  janitor::clean_names()

```

# Demographics

```{r}

dat <- data_raw |> 
  select(-c(topic_type, data_source, data_value_unit, data_value_type, topic_type, 
            response, data_value_footnote_symbol)) |> 
  filter(topic_desc == "Cigarette Use (Adults)",
         measure_desc == "Current Smoking",
         location_desc == "Alaska",
         education == "All Grades" 
         ) 

distinct(dat, education)

```

# Exclusions / data quality

## AMP

```{r}

```


# Define master exclusions

```{r}

# create a master exclude_participant variable
data_processed <- data_processed_before_exclusions |>
  mutate(exclude_participant = case_when(tolower(age) == "test" ~ "exclude",
                                         tolower(gender) == "test" ~ "exclude",
                                         is.na(mean_evaluation) ~ "exclude",
                                         # in this case we will exclude participants with missing demographics data or outcomes measures data. 
                                         # Note that "list-wise exclusions" like this aren't always justified, as missingness often isn't at random. 
                                         # How to treat missing data is a  whole area of work in itself, which we wont cover here.
                                         is.na(age) ~ "exclude", 
                                         is.na(gender) ~ "exclude",
                                         exclude_amp_performance == "exclude" ~ "exclude",
                                         exclude_duplicate_data == "exclude" ~ "exclude",
                                         exclude_amp_completeness == "exclude" ~ "exclude", 
                                         TRUE ~ "include"))

```

# Write to disk

```{r}

# in case this dir doesn't exist, create it
dir.create("../data/processed/")

# save data to disk in that dir
write_csv(data_processed, "../data/processed/data_processed.csv")

```

# Create codebook template for the processed data


```{r}
if(!file.exists("../data/processed/data_processed_codebook.xlsx")){
  # convert the column names to a df
  codebook_template <- data.frame(variable = colnames(data_processed)) |>
    mutate(explanation = NA)
  # write to disk as an excel file
  openxlsx::write.xlsx(codebook_template, file = "../data/processed/data_processed_codebook.xlsx")
}
```

Note that there are other ways of automatically creating more elaborate codebooks from your datasets. These often contain information about min/max/mean/SD, distribution, etc. For example:

- Ruben Arslan's {codebook}
  - [R package with How-Tos](https://rubenarslan.github.io/codebook/)
  - [Tutorial](https://rubenarslan.github.io/codebook/articles/codebook_tutorial.html)
  - [Article](https://journals.sagepub.com/doi/full/10.1177/2515245919838783)
- Petersen & Ekstrøm's {dataReporter}
  - [Article](https://www.jstatsoft.org/article/view/v090i06)
  - [Blog](https://sandsynligvis.dk/2017/08/21/datamaid-your-personal-assistant-for-cleaning-up-the-data-cleaning-process/)
  - [R package](https://cran.r-project.org/web/packages/dataReporter/index.html)

# Session info

```{r}

sessionInfo()

```


