---
title: "Evaluations of positive and negative stimuli using the Affective Misattribution Procedure (AMP) and self-reports"
subtitle: "Analysis"
author: "Template: Ian Hussey; content: [Student name]"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r, include=FALSE}

# set knit options
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)

# disable scientific notation
options(scipen = 999) 

```

# Dependencies

```{r}

library(tidyverse)
library(knitr)
library(kableExtra)
library(janitor)
library(scales)

```

# Data

Load the processed data and apply the global exclusions.

```{r}

data_processed <- read_csv("../data/processed/data_processed.csv")

data_processed_after_exclusions <- data_processed |>
  filter(exclude_participant == "include")

```

# Sample descriptives

## Sample size before exclusions

```{r}

data_processed |>
  count(name = "n") |>
  kable() |>
  add_header_above(header = c("Whole sample" = 1)) |> # note that you can add header rows to tables like this. The "1" indicates the number of columns the header should span. The sum of these numbers must equal the number of columns or you'll get an error.
  kable_classic(full_width = FALSE)

```

## Sample size after exclusions

Sample used in subsequent analyses

```{r}

data_processed_after_exclusions |>
  count(name = "n") |>
  kable() |>
  add_header_above(header = c("For analysis" = 1)) |>
  kable_classic(full_width = FALSE)

```

## Age

```{r}

data_processed_after_exclusions |>
  mutate(age = as.numeric(age)) |>
  summarise(Mean = mean(age, na.rm = TRUE),
            SD = sd(age, na.rm = TRUE)) |>
  mutate_all(.funs = janitor::round_half_up, digits = 1) |>
  kable() |>
  add_header_above(header = c("Age" = 2)) |>
  kable_classic(full_width = FALSE)

```

## Gender

```{r}

data_processed_after_exclusions |> 
  rename(Gender = gender) |>
  group_by(Gender) |> 
  summarise(n = n()) |> 
  mutate(Percent = paste0(round_half_up((n / sum(n)) * 100, 1), "%")) |>
  mutate(Gender = stringr::str_to_sentence(Gender)) |> # Change the case of the Gender variable so that it prints nicely
  kable() |>
  kable_classic(full_width = FALSE)

```

# Descriptives

Descriptive statistics and plots of the measures (excluding the demographics variables)

## Self-reported evaluations

### Descriptive stats

```{r}

# overall self-reported evaluations
dat_mean_ratings <- data_processed_after_exclusions |>
  summarise(Mean = mean(mean_evaluation, na.rm = TRUE),
            SD = sd(mean_evaluation, na.rm = TRUE),
            n = n()) |>
  mutate(group = "Full sample")

# self-reported evaluations by gender category
dat_mean_ratings_by_gender <- data_processed_after_exclusions |>
  group_by(group = gender) |>
  summarise(Mean = mean(mean_evaluation, na.rm = TRUE),
            SD = sd(mean_evaluation, na.rm = TRUE),
            n = n())

# combine both into one table
bind_rows(dat_mean_ratings,
          dat_mean_ratings_by_gender) |>
  select(Subset = group, Mean, SD, n) |> # select variables of interest, and rename one 
  mutate(Subset = stringr::str_to_sentence(Subset)) |> # Change the case of the Subset variable so that it prints nicely
  mutate_if(is.numeric, round_half_up, digits = 2) |>
  kable() |>
  add_header_above(header = c(" " = 1, "Self-reported evaluations" = 3)) |>
  kable_classic(full_width = FALSE)

```

### Descriptive plot

```{r}

ggplot(data_processed_after_exclusions, aes(x = mean_evaluation)) +
  geom_histogram(binwidth = 1,
                 boundary = 0,
                 fill = viridis_pal(begin = 0.45, option = "mako")(1), 
                 color = viridis_pal(begin = 0.30, option = "mako")(1)) + 
  xlab("Mean self-reported evaluation") +
  ylab("Frequency") +
  theme_linedraw() +
  scale_x_continuous(breaks = pretty_breaks(n = 7)) +
  coord_cartesian(xlim = c(1, 7)) +
  theme(panel.grid.minor = element_blank())

```

## AMP evaluations

### Descriptive stats

\TODO add table of means, SDs, Ns

```{r}
# overall self-reported evaluations
dat_amp_ratings <- data_processed_after_exclusions |>
  summarise(Mean = mean(AMP_score, na.rm = TRUE),
            SD = sd(AMP_score, na.rm = TRUE),
            n = n()) |>
  mutate(group = "Full sample")

# self-reported evaluations by gender category
dat_amp_ratings_by_gender <- data_processed_after_exclusions |>
  group_by(group = gender) |>
  summarise(Mean = mean(AMP_score, na.rm = TRUE),
            SD = sd(AMP_score, na.rm = TRUE),
            n = n())

# combine both into one table
bind_rows(dat_amp_ratings,
          dat_amp_ratings_by_gender) |>
  select(Subset = group, Mean, SD, n) |> # select variables of interest, and rename one 
  mutate(Subset = stringr::str_to_sentence(Subset)) |> # Change the case of the Subset variable so that it prints nicely
  mutate_if(is.numeric, round_half_up, digits = 2) |>
  kable() |>
  add_header_above(header = c(" " = 1, "Self-reported evaluations" = 3)) |>
  kable_classic(full_width = FALSE)

```

### Descriptive plots

```{r}

ggplot(data_processed_after_exclusions, aes(x = AMP_score)) +
  geom_histogram(binwidth = 0.05,
                 boundary = 0,
                 fill = viridis_pal(begin = 0.45, option = "mako")(1), 
                 color = viridis_pal(begin = 0.30, option = "mako")(1)) + 
  xlab("AMP score") +
  ylab("Frequency") +
  theme_linedraw() +
  scale_x_continuous(breaks = pretty_breaks(n = 10))

```

# Analyses & hypothesis tests

## Self-reported evaluations are correlated with evaluations on the AMP

### Plot

```{r}

ggplot(data_processed_after_exclusions, 
       aes(x = AMP_score,
           y = mean_evaluation)) +
  geom_jitter(color = viridis_pal(begin = 0.45, option = "mako")(1),
              alpha = 0.5) +
  geom_smooth(method = "lm",
              color = viridis_pal(begin = 0.45, option = "mako")(1)) +
  xlab("AMP score") +
  ylab("Mean self-reported evaluation") +
  theme_linedraw() 

ggplot(data_processed_after_exclusions, 
       aes(y = AMP_score,
           x = mean_evaluation)) +
  geom_jitter(color = viridis_pal(begin = 0.45, option = "mako")(1),
              alpha = 0.5) +
  geom_smooth(method = "lm",
              color = viridis_pal(begin = 0.45, option = "mako")(1)) +
  ylab("AMP score") +
  xlab("Mean self-reported evaluation") +
  theme_linedraw() 

ggplot(data_processed_after_exclusions, 
       aes(x = AMP_score,
           y = mean_evaluation)) +
  geom_jitter(color = viridis_pal(begin = 0.45, option = "mako")(1),
              alpha = 0.5) +
  xlab("AMP score") +
  ylab("Mean self-reported evaluation") +
  theme_linedraw() 

```

More complex plots:

#### Axial histograms

Scatter plots with axial histograms using ggExtra: https://cran.r-project.org/web/packages/ggExtra/vignettes/ggExtra.html

\TODO add axial histograms to a scatter plot. Split both the scatter plot and the histograms by gender.

```{r}

data_female <- subset(data_processed_after_exclusions, gender == "female")
data_male <- subset(data_processed_after_exclusions, gender == "male")

p_female <- ggplot(data_female, 
       aes(x = AMP_score,
           y = mean_evaluation)) +
  geom_jitter(color = viridis_pal(begin = 0.45, option = "mako")(1),
              alpha = 0.5)

p_male <- ggplot(data_female, 
       aes(x = AMP_score,
           y = mean_evaluation)) +
  geom_jitter(color = viridis_pal(begin = 0.45, option = "mako")(1),
              alpha = 0.5)

p_male_hist <- ggExtra::ggMarginal(p_male, type = "histogram")
p_female_hist <- ggExtra::ggMarginal(p_female, type = "histogram")

gridExtra::grid.arrange(p_male_hist, p_female_hist, ncol = 1)
```
```{r}
p1 <- ggplot(data_processed_after_exclusions, 
       aes(x = AMP_score,
           y = mean_evaluation, 
           colour = gender)) +
    geom_jitter(alpha = 0.5) +
  xlab("AMP score") +
  ylab("Mean self-reported evaluation")

  ggExtra::ggMarginal(p1, groupColour = TRUE, groupFill = TRUE)


```



#### Labelled points

Label points using ggrepel: https://cran.r-project.org/web/packages/ggrepel/vignettes/ggrepel.html

\TODO Label the points in a scatter plot using their participant codes. Label only the participants with more extreme scores.

```{r}
library(ggrepel)

# Berechnen der 95. Perzentile f체r AMP_score und mean_evaluation
quantile_95_mean_evaluation <- quantile(data_processed_after_exclusions$mean_evaluation, 0.95)


# Filtern der Daten f체r Werte im 95. Perzentil
extreme_values <- data_processed_after_exclusions |> 
  filter(mean_evaluation >= quantile_95_mean_evaluation)

ggplot(data_processed_after_exclusions, 
          aes(x = AMP_score,
              y = mean_evaluation,
              label = subject)) +
geom_jitter(color = viridis_pal(begin = 0.45, option = "mako")(1),
            alpha = 0.5) + 
  geom_text_repel(data = extreme_values,
                  aes(label = subject),
                  box.padding = 0.35,
                  point.padding = 0.5,
                  segment.color = 'grey50')

```

#### Magnify areas

Magnify areas of your plot with ggmagnify: https://hughjonesd.github.io/ggmagnify/

\TODO Magnify an area of one of your scatter plots, eg where there are a lot of data points in a small area.

```{r}
library(ggmagnify)

p2 <- ggplot(data_processed_after_exclusions, 
       aes(x = AMP_score,
           y = mean_evaluation, 
           colour = gender)) +
  geom_jitter(alpha = 0.5) +
  geom_magnify(shape = "rect",
               from = c(xmin = 0.4, xmax = 0.6, ymin = 0.8, ymax = 1.3 ),
               to = c(xmin = 0.75, xmax = 1, ymin = 4, ymax = 5.8 )) +
  theme_minimal()

p2

```

### Test

\TODO run an appropriate test. Below the output, interpret the results: write a few sentences that report and interpret the results following APA reporting style.

```{r}

cor.test(data_processed_after_exclusions$AMP_score, data_processed_after_exclusions$mean_evaluation, 
         method = "pearson")


```
there is a small negativ correlation between the AMP-Score and the evaluation mean. This correlation is not significant. 


### Test

\TODO 

## Self-reported evalautions differ between men and women

### Plot

\TODO split histogram, split violin plot, raincloud plot, etc.

```{r}

ggplot(data_processed_after_exclusions, 
       aes(x = mean_evaluation, 
           fill = gender)) +
  geom_histogram() +
  facet_grid(~ gender) +
  theme_minimal()


ggplot(data_processed_after_exclusions, 
       aes(y = mean_evaluation,
           x = gender,
           fill = gender, 
           color = gender)) +
  geom_violin(alpha = 0.5, size = 0.5) +
  theme_minimal()

```

### Test

\TODO run an appropriate test. Below the output, interpret the results: write a few sentences that report and interpret the results following APA reporting style.

```{r}

filtered_data <- data_processed_after_exclusions |> 
  filter(gender %in% c("male", "female"))
  
ttest_mean.eval_gender <- t.test(mean_evaluation ~ gender, data = filtered_data)

```
In der Analyse der Mittelwertsunterschiede wurde die Gruppe "non-binary" ausgeschlossen, da sie mit $n=2$ zu wenig Datenpunkte aufweist. F체r die zwei Geschlechter m채nnlich und weiblich zeigt sich auf einem Level von ${p=0.05}$ und einem Wert von $p =$ `r round(ttest_mean.eval_gender$p.value, 2)`, dass die Unterschiede in der Mean Evaluation zwischen den Geschlechtern signifikant ist.  


## Evaluations on the Affect Misattribution Procedure differ between men and women

### Plot

\TODO split histogram, split violin plot, raincloud plot, etc.

This time, vary the labeling and order of the legend, e.g., capitalise "Men" and "Women", and know how to change the order of the factors.

```{r}

ggplot(data_processed_after_exclusions, 
       aes(y = mean_evaluation,
           x = gender,
           fill = gender, 
           color = gender)) +
  geom_violin(alpha = 0.5, size = 0.5) +
  scale_x_discrete(limits = c("male", "female", "nonbinary")) +
  scale_x_discrete(limits = c("male", "female", "nonbinary"), 
                   labels = c("Male", "Female", "Non-Binary")) +
  scale_fill_discrete(limits = c("male", "female", "nonbinary"),
                      name = "Gender", 
                      labels=c("Male", "Female", "Non-Binary")) +
  scale_color_discrete(limits = c("male", "female", "nonbinary"),
                       name = "Gender", labels=c("Male", "Female", "Non-Binary")) +
  labs(x = "Gender", y = "Mean Evaluation") +
  theme_minimal()
```
```{r}
ggplot(data_processed_after_exclusions, aes(x = AMP_score, fill = gender)) +
  geom_histogram(binwidth = 0.15, color = "black", alpha = 0.7) +
  facet_grid(. ~ gender) +
  scale_fill_manual(values = c("female" = "red", "male" = "blue", "nonbinary" = "yellow"),
                    labels = c("Female", "Male", "Non-binary")) +
  labs(x = "AMP score", y = "Mean self-reported evaluation", fill = "Gender") +
  theme_minimal()

```

### Test

\TODO run an appropriate test. Below the output, print an interpretation of the results generated by the 'easystats' package [report](https://easystats.github.io/report/). I.e., use `report::report()`.

```{r}
library(report)

filtered_data <- data_processed_after_exclusions |> 
  filter(gender %in% c("male", "female"))
  
ttest_mean.eval_gender <- t.test(mean_evaluation ~ gender, data = filtered_data)
report(ttest_mean.eval_gender)

```

## Combining plots

Combine plots using patchwork: https://patchwork.data-imaginist.com/

\TODO Combine at least three of the above plots into one.

```{r}
library(patchwork)

p1 <- ggplot(data_processed_after_exclusions, 
       aes(x = AMP_score,
           y = mean_evaluation)) +
  geom_jitter(color = viridis_pal(begin = 0.45, option = "mako")(1),
              alpha = 0.5) +
  xlab("AMP score") +
  ylab("Mean self-reported evaluation") +
  theme_linedraw() 

p2 <- ggplot(data_processed_after_exclusions, 
       aes(y = mean_evaluation,
           x = gender,
           fill = gender, 
           color = gender)) +
  geom_violin(alpha = 0.5, size = 0.5) +
  scale_x_discrete(limits = c("male", "female", "nonbinary")) +
  scale_x_discrete(limits = c("male", "female", "nonbinary"), 
                   labels = c("Male", "Female", "Non-Binary")) +
  scale_fill_discrete(limits = c("male", "female", "nonbinary"),
                      name = "Gender", labels=c("Male", "Female", "Non-Binary")) +
  scale_color_discrete(limits = c("male", "female", "nonbinary"),
                       name = "Gender", labels=c("Male", "Female", "Non-Binary")) +
  labs(x = "Gender", y = "Mean Evaluation") +
  theme_minimal()

p3 <- ggplot(data_processed_after_exclusions,
             aes(x = AMP_score,
             y = mean_evaluation, 
             colour = gender)) +
    geom_jitter(alpha = 0.5) +
    geom_magnify(shape = "rect",
                 from = c(xmin = 0.4, xmax = 0.6, ymin = 0.8, ymax = 1.3 ),
                 to = c(xmin = 0.75, xmax = 1, ymin = 4, ymax = 5.8 )) +
    theme_minimal()

p1 + p2 + p3 +
  plot_layout(widths = c(2, 2))

```

## Saving plots

Save plots to disk with `ggsave()`

\TODO Save the above combined plot to disk as both .png and .pdf. Ensure the png has at least 300dpi resolution.

```{r}

ggsave("combined_plots.png", dpi = 300)
ggsave("combined_plots.pdf", dpi = 300)

```

# Session info

```{r}

sessionInfo()

```



